Critical assessment of Linear-Response Excitations in Molecular system\\ or \\ On the Localization of the Linear-Response excitations in Open Systems

### 2017-12-23 LRDraft.tex
update

Let us extend the the former analysis to self-consistent systems, \emph{i.e.} to the case in which the unperturbed Hamiltonian is itself a function of the density.
In this case the perturbing field $\op\Phi$ has also the effect of adding a further perturbation operator to the perturbing Hamiltonian. In the linear response framework, 
this latter quantity can be expressed as the functional derivatives of $\hnot$ with respect to the density kernel (evaluated at $\dm=\dmnot$) times the value of the
induced density modification, that is:
\be\label{V'tspacedef1}
\op V'(t)= 
\int \dd \r \dd \r' \frac{\delta\hnot}{\delta \rho_0(\r,\r')}
\trace{\dm'(t) \ket{\r} \bra{\r'}} \;.
\ee
The Liouville equation for self-consistent systems can be written by adding the contribution of $\op V'$ to $\op\Phi$ in \eqref{LiouvilleRhoDef2}. 
It reads:
\be\lb{LiouvilleRhoSelfDef1}
\ii \d_t \dm'(t) = \Liouvnot \dm'(t) + \commutator{\op V'(t)}{\dmnot} + \commutator{\op \Phi(t)}{\dmnot} \;.
\ee
We point out that the response density operator obeys a self-consistent equation, since the  induced perturbing field $\op V'$ is itself a function of $\dm'$. 
It is possible to formally recast \eqref{LiouvilleRhoSelfDef1} in the same form of \eqref{LiouvilleRhoDef2} by introducing the \emph{coupling}
superoperator:
\be
\coupl = 
\int \dd \r \dd \r' \commutator{\frac{\delta\hnot}{\delta \rho_0(\r,\r')}}{\dmnot}
\trace{\cdot \ket{\r} \bra{\r'}} \;,
\ee
in this way the first commutator in \eqref{LiouvilleRhoSelfDef1} can be written as the formal action of the coupling superoperator to $\dm'$ and \eqref{LiouvilleRhoSelfDef1}
can be written as:
\be\lb{LiouvilleRhoSelfDef2}
\ii \d_t \dm'(t) = \Liouv \dm'(t) + \commutator{\op \Phi(t)}{\dmnot} \;, \qq
\Liouv = \Liouvnot + \coupl \;.
\ee

@@

We reformulate the Liouville equation \eqref{LiouvilleRhoSelfDef2} in the frequency domain. In order to achieve this task it is
preferable to directly compute the action of the Liouville operator on the Fourier transform of the response density, since in this
way some subtleties that concerns the analytic structure in the $\omega$ complex plane can be properly taken into account.
We first define the response density in the frequency domain that, staring from \eqref{rhoTimeEvolutionDef3}, is given by:
\be\lb{rhoPrimeOmegaDef1}
\dm'(\omega) \equiv \intover{t} e^{\ii \omega t} \dm'(t)
= \ii \commutator{\dmnot}{\op A(\omega)} \;,
\ee
where $\op A(\omega)$ is The Fourier transform of \eqref{AopDef1} and reads:
\be \lb{AomegaDef1}
\op A(\omega) = \frac{1}{2\pi} \int_{-\infty}^\infty  \dd \omega'
\op G_0^+(\omega') \op \Phi(\omega) \op G_0^-(\omega'-\omega)\;.
\ee
The action of $\Liouvnot$ on $\dm'(\omega)$ follows from \eqref{L0sopDef1}:
\be\lb{LiouvillianRho'omegaDef1}
\Liouvnot \dm'(\omega) = \ii \commutator{\hnot}{\commutator{\dmnot}{\op A(\omega)}} = 
\ii \commutator{\dmnot}{\commutator{\hnot}{\op A(\omega)}} \;,
\ee
and the commutator between the unperturbed Hamiltonian and $\op A(\omega)$ can be directly computed starting from \eqref{AomegaDef1}: 
\be
\commutator{\hnot }{\op A(\omega)} = \bbomega \op A(\omega) - \ii \op \Phi(\omega)\;, \nn
\ee
where $\bbomega \equiv \omega + 2\ii\eta$, so that:
\be\lb{LiouvillianRho'omegaDef2}
\Liouvnot \dm'(\omega) =  \bbomega\dm'(\omega) + \commutator{\dmnot}{\op\Phi(\omega)}\;.
\ee
The coupling superoperator has the same formal action on the response density both in the time and in the frequency domain, so the complete
Liouville equation in $\omega$ space can be written as:
\be\lb{LiouvillianRho'omegaDef3}
\left(\bbomega - \Liouv\right) \dm'(\omega) =  \commutator{\op\Phi(\omega)}{\dmnot}\;.
\ee

@@

We split the complete
set of eigenstates of $\hnot$ in two sets. The first one is composed by the eigenvectors of $\hnot$ in the subspace of the occupied orbitals. The second 
one is composed by the remaining eigenvectors, both in the bound and in the continuum region, and its elements are denoted as $\ket{\psi(\alpha)}$, where $\alpha$ 
is a multi-index in the sense described in \ref{BasicDef}. The elements of these sets satisfy the following orthonormality relation: 
\be\lb{orthonormalityDef1}
\brket{\psi(\alpha)}{\psi_p} = \brket{\psi_p}{\psi(\alpha)} = 0 \;, \qq \brket{\psi_p}{\psi_q} = \delta_{pq} \;, \qq
\brket{\psi(\alpha)}{\psi(\beta)} = \delta(\alpha-\beta) \;.
\ee
Obviously, the direct sum of these two sets coincides with the complete set of eigenstates of $\hnot$ and thus 
constitute a complete basis. 
A generic operator can be decomposed in this basis and its formal expansion contains four distinct contribution of the form $\ketbra{\psi_p}{\psi_q}$, 
$\ketbra{\psi(\alpha)}{\psi(\beta)}$, $\ketbra{\psi(\alpha)}{\psi_p}$, $\ketbra{\psi_p}{\psi(\alpha)}$. However an inspection of equations \eqref{rhoPrimeOmegaDef1}
and \eqref{LiouvillianRho'omegaDef3} evidences that the relevant operators in the contest of the linear response are expressed as a commutator with $\dmnot$. 
For this class of operators the only non vanishing contribution comes from the \emph{transition operators}:
\be
\excite{p}{\alpha} \equiv \ketbra{\psi(\alpha)}{\psi_p}\;, \qq
\decay{\alpha}{p} \equiv \ketbra{\psi_p}{\psi(\alpha)} \;.
\ee
The orthonormality conditions \eqref{orthonormalityDef1} imply that:
\be\lb{transitionOpOrthonormalityDef1}
\trace{\excite{p}{\alpha}\excite{q}{\beta}} = \trace{\decay{\alpha}{p}\decay{\beta}{q}} = 0 \;, \qq
\trace{\excite{p}{\alpha}\decay{\beta}{q}} = \delta_{pq}\delta(\alpha-\beta) \;,
\ee
so that the expansion of a suitable operator $\op O$ in the basis of transitions reads:
\be\lb{OperatorTransBasisExpDef1}
\op O = \sum_p\sint \dd \alpha\left(
\excite{p}{\alpha}\trace{\decay{\alpha}{p}\op O} + \decay{\alpha}{p}\trace{\excite{p}{\alpha}\op O}
\right)
\ee
In order to recast this expansion in a simpler expression it turns out to be useful to define the rank-2 vectors of the transition basis: 
\be
\sbra{p\alpha}{\cdot} \equiv 
\mat{ \excite{p}{\alpha} \;, & \decay{\alpha}{p} }\;,
\ee
as well as a different representation for the trace between the operators and the above mentioned transitions:
\be\lb{OperatorTransitionCoeffDef1}
\sket{p\alpha}{\op O}  \equiv \mat{ \trace{\decay{\alpha}{p}\op O } \\ 
\trace{\excite{p}{\alpha}\op O}} \; .
\ee
Thanks to this formalism we have that equation \eqref{OperatorTransBasisExpDef1} can be expressed as:
\be\lb{OperatorTransBasisExpDef2}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\cdot}\sket{p\alpha}{\op O} = 
\dbraket{\cdot}{\op O} \;,
\ee
where the last equality implicitly define the index-less ``big-shaped'' vectors and the sum over unwritten indexes is understood.

\subsubsection{Spectral decompositionof the response density in the transitions basis}

The Liouville equation \eqref{LiouvillianRho'omegaDef3} can be consistently written in the transition basis. Indeed, its r.h.s is explicitly written
as a commutator with $\dmnot$, while an inspection of the formal definition of the response density and of the Liouville superoperator evidences that
they both contain a commutator with $\dmnot$. Written in the basis of transitions Liouville equation becomes a linear matrix equation for the rank-2 
vectors of the coefficients \eqref{OperatorTransitionCoeffDef1} of the operators in its left and right side:
\be
\omega\mat{ \trace{\decay{\alpha}{p}\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\dm'(\omega)}} - 
\mat{ \trace{\decay{\alpha}{p}\Liouv\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\Liouv\dm'(\omega)}} =
\mat{ \trace{\decay{\alpha}{p}\commutator{\op\Phi(\omega)}{\dmnot}} \\ \trace{\excite{p}{\alpha}\commutator{\op\Phi(\omega)}{\dmnot}}}
\;.\nn
\ee
The second addend in the l.h.s can be further analyzed by writing the transition basis expansion of $\dm'$ inside the traces. Performing this procedure
we can recast this quantity as the action of a 2-dimensional matrix on the rank-2 vector of the response density. So we have: 
\be\lb{LiouvilleTransBasisDef1}
\sum_q\sint\dd\beta \left[\omega\delta_q^p\delta(\alpha-\beta)- \optr{L}^{p \alpha}_{q\beta} \right]\sket{q\beta}{\dm'(\omega)} =
\sket{p\alpha}{\commutator{\op\Phi(\omega)}{\dmnot}} \;.
\ee
with:
\be\lb{LiouvilleMatrixTransBasisDef1}
\optr{L}^{p \alpha}_{q\beta} \equiv \mat{\trace{\decay{\alpha}{p}\, \Liouv\, \excite{q}{\beta}} &
\trace{\decay{\alpha}{p}\, \Liouv\, \decay{\beta}{q}} \\
\trace{\excite{p}{\alpha}\, \Liouv\, \excite{q}{\beta}} &
\trace{\excite{p}{\alpha}\, \Liouv\, \decay{\beta}{q}}} \;.
\ee
This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted:
\be
\sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
\frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
\sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
that can be compactly rewritten by using the index-less notation:
\be
\sket{p\alpha}{\dm'(\omega)} = 
\sint\dd E \sket{p\alpha}{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee

@@

We introduce the explicit representation of the eigenvectors op $\optr{L}$:
\be
\sket{p\alpha}{E} \equiv \mat{X_E^{p\alpha} \\ Y_E^{p\alpha}} \;, \qq
\sbra{p\alpha}{\tilde E} \equiv \mat{\tilde X_{E\,p\alpha} & \tilde Y_{E\,p\alpha}} \;,
\ee
and on the basis of these elements we define the \emph{excitation operators}:
\be\lb{ExctitationOperatorsDef1}
\op E = 
\sum_p \sint \dd \alpha \left(\excite{p}{\alpha} X_E^{p\alpha} + \decay{\alpha}{p} Y_E^{p\alpha}\right) 
\;, \qq
\op{\tilde E} = \sum_p \sint \dd \alpha \left(\decay{\alpha}{p} \tilde X_{E\,p\alpha} + 
\excite{p}{\alpha}\tilde Y_{E\,p\alpha}  \right)\;.
\ee
Excitation operators are the right and left eigenvectors of the Liouville superoperator, as directly follows from their
definition. Indeed for $\op E$ we have:
\begin{align}
\sop{\mathcal L} \op E & = \sum_p \sint \dd \alpha \left(\trace{\decay{\alpha}{p} \sop{\mathcal L} \op E} \excite{p}{\alpha} + 
\decay{\alpha}{p} \trace{ \excite{p}{\alpha} \sop{\mathcal L} \op E}  \right) = \nn \\
&= \sum_p \sint  \dd \alpha \sbra{p\alpha}{\cdot} \sum_q \sint  \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q\beta}{E} 
= E \sum_p \sint \dd \alpha \sbra{p\alpha}{\cdot} \sket{p\alpha}{E} = E \, \op E \;,	
\end{align}
whereas the left action on $\op{\tilde E}$ provides:
\begin{align}
 \op{\tilde E} \sop{\mathcal L} &= E \op{\tilde E}
\end{align}

>>>>

Let us extend the the former analysis to self-consistent systems, \emph{i.e.} to the case in which the unperturbed Hamiltonian is itself a function of the density.
In this case the perturbing field $\op\Phi$ has also the effect of adding a further perturbation operator to the perturbing Hamiltonian. In the linear response framework, 
this latter quantity can be expressed as the functional derivatives of $\hnot$ with respect to the density kernel (evaluated at $\dm=\dmnot$) times the value of the
induced density modification, that is:
\be\label{Vptspacedef1}
\op V'(t)= 
\int \dd \r \dd \r' \frac{\delta\hnot}{\delta \rho_0(\r,\r')}
\trace{\dm'(t) \ket{\r} \bra{\r'}} \;.
\ee
The Liouville equation for self-consistent systems can be written by adding the contribution of $\op V'$ to $\op\Phi$ in \eqref{LiouvilleRhoDef2}. 
It reads:
\be\lb{LiouvilleRhoSelfDef1}
\ii \d_t \dm'(t) = \Liouvnot \dm'(t) + \commutator{\op V'(t)}{\dmnot} + \commutator{\op \Phi(t)}{\dmnot} \;.
\ee
We point out that the response density operator obeys a self-consistent equation, since the  induced perturbing field $\op V'$ is itself a function of $\dm'$. 
It is possible to formally recast \eqref{LiouvilleRhoSelfDef1} in the same form of \eqref{LiouvilleRhoDef2} by introducing the \emph{coupling}
superoperator:
\be
\coupl = 
\int \dd \r \dd \r' \commutator{\frac{\delta\hnot}{\delta \rho_0(\r,\r')}}{\dmnot}
\trace{\cdot \ket{\r} \bra{\r'}} \;,
\ee
in this way the first commutator in \eqref{LiouvilleRhoSelfDef1} can be written as the formal action of the coupling superoperator to $\dm'$ and \eqref{LiouvilleRhoSelfDef1}
can be written as:
\be\lb{LiouvilleRhoSelfDef2}
\ii \d_t \dm'(t) = \Liouv \dm'(t) + \commutator{\op \Phi(t)}{\dmnot} \;, \qq
\Liouv = \Liouvnot + \coupl \;.
\ee

@@

We reformulate the Liouville equation \eqref{LiouvilleRhoSelfDef2} in the frequency domain. In order to achieve this task it is
preferable to directly compute the action of the Liouville operator on the Fourier transform of the response density, since in this
way some subtleties that concerns the analytic structure in the $\omega$ complex plane can be properly taken into account.
We first define the response density in the frequency domain that, staring from \eqref{rhoTimeEvolutionDef3}, is given by:
\be\lb{rhoPrimeOmegaDef1}
\dm'(\omega) \equiv \intover{t} e^{\ii \omega t} \dm'(t)
= \ii \commutator{\dmnot}{\op A(\omega)} \;,
\ee
where $\op A(\omega)$ is The Fourier transform of \eqref{AopDef1} and reads:
\be \lb{AomegaDef1}
\op A(\omega) = \frac{1}{2\pi} \int_{-\infty}^\infty  \dd \omega'
\op G_0^+(\omega') \op \Phi(\omega) \op G_0^-(\omega'-\omega)\;.
\ee
The action of $\Liouvnot$ on $\dm'(\omega)$ follows from \eqref{L0sopDef1}:
\be\lb{LiouvillianRhopomegaDef1}
\Liouvnot \dm'(\omega) = \ii \commutator{\hnot}{\commutator{\dmnot}{\op A(\omega)}} = 
\ii \commutator{\dmnot}{\commutator{\hnot}{\op A(\omega)}} \;,
\ee
and the commutator between the unperturbed Hamiltonian and $\op A(\omega)$ can be directly computed starting from \eqref{AomegaDef1}: 
\be
\commutator{\hnot }{\op A(\omega)} = \bbomega \op A(\omega) - \ii \op \Phi(\omega)\;, \nn
\ee
where $\bbomega \equiv \omega + 2\ii\eta$, so that:
\be\lb{LiouvillianRhopomegaDef2}
\Liouvnot \dm'(\omega) =  \bbomega\dm'(\omega) + \commutator{\dmnot}{\op\Phi(\omega)}\;.
\ee
The coupling superoperator has the same formal action on the response density both in the time and in the frequency domain, so the complete
Liouville equation in $\omega$ space can be written as:
\be\lb{LiouvillianRhopomegaDef3}
\left(\bbomega - \Liouv\right) \dm'(\omega) =  \commutator{\op\Phi(\omega)}{\dmnot}\;.
\ee

@@

We split the complete
set of eigenstates of $\hnot$ in two sets. The first one is composed by the eigenvectors of $\hnot$ in the subspace of the occupied orbitals. The second 
one is composed by the remaining eigenvectors, both in the bound and in the continuum region, and its elements are denoted as $\ket{\psi(\alpha)}$, where $\alpha$ 
is a multi-index in the sense described in \ref{BasicDef}. The elements of these sets satisfy the following orthonormality relation: 
\be\lb{orthonormalityDef1}
\brket{\psi(\alpha)}{\psi_p} = \brket{\psi_p}{\psi(\alpha)} = 0 \;, \qq \brket{\psi_p}{\psi_q} = \delta_{pq} \;, \qq
\brket{\psi(\alpha)}{\psi(\beta)} = \delta(\alpha-\beta) \;.
\ee
Obviously, the direct sum of these two sets coincides with the complete set of eigenstates of $\hnot$ and thus 
constitute a complete basis. 
A generic operator can be decomposed in this basis and its formal expansion contains four distinct contribution of the form $\ketbra{\psi_p}{\psi_q}$, 
$\ketbra{\psi(\alpha)}{\psi(\beta)}$, $\ketbra{\psi(\alpha)}{\psi_p}$, $\ketbra{\psi_p}{\psi(\alpha)}$. However an inspection of equations \eqref{rhoPrimeOmegaDef1}
and \eqref{LiouvillianRhopomegaDef3} evidences that the relevant operators in the contest of the linear response are expressed as a commutator with $\dmnot$. 
For this class of operators the only non vanishing contribution comes from the \emph{transition operators}:
\be
\excite{p}{\alpha} \equiv \ketbra{\psi(\alpha)}{\psi_p}\;, \qq
\decay{\alpha}{p} \equiv \ketbra{\psi_p}{\psi(\alpha)} \;.
\ee
The orthonormality conditions \eqref{orthonormalityDef1} imply that:
\be\lb{transitionOpOrthonormalityDef1}
\trace{\excite{p}{\alpha}\excite{q}{\beta}} = \trace{\decay{\alpha}{p}\decay{\beta}{q}} = 0 \;, \qq
\trace{\excite{p}{\alpha}\decay{\beta}{q}} = \delta_{pq}\delta(\alpha-\beta) \;,
\ee
so that the expansion of a suitable operator $\op O$ in the basis of transitions reads:
\be\lb{OperatorTransBasisExpDef1}
\op O = \sum_p\sint \dd \alpha\left(
\excite{p}{\alpha}\trace{\decay{\alpha}{p}\op O} + \decay{\alpha}{p}\trace{\excite{p}{\alpha}\op O}
\right) \;.
\ee
In order to recast this expansion in a simpler expression it turns out to be useful to define the rank-2 row vectors of the transition basis: 
\be
\sbra{p\alpha}{\cdot} \equiv 
\mat{ \excite{p}{\alpha} \;, & \decay{\alpha}{p} }\;,
\ee
and we also introduce a vectorial notation to deal with the coefficients of the expansion of an operator in the transition basis, that is: 
\be\lb{OperatorTransitionCoeffDef1}
\sket{p\alpha}{\op O}  \equiv \mat{ \trace{\decay{\alpha}{p}\op O } \\ 
\trace{\excite{p}{\alpha}\op O}} \; .
\ee
Thanks to this formalism we have that equation \eqref{OperatorTransBasisExpDef1} can be expressed as:
\be\lb{OperatorTransBasisExpDef2}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\cdot}\sket{p\alpha}{\op O} = 
\dbraket{\cdot}{\op O} \;,
\ee
where the last equality implicitly define the ulterior compact notation of the ``big-shaped'' vectors in which the summed indexes are omitted.  
In the same fashion it will be also useful for the subsequent analysis to introduce a ``conjugate'' notation, in which the basis operators are expressed
as column vector and the coefficients of the expanded operator as row ones, that is: 
\be\lb{TransitionBasisConjugateDef1}
\sket{p\alpha}{\cdot} \equiv 
\mat{\decay{\alpha}{p} \\  \excite{p}{\alpha}} \;, \qq
\sbra{p\alpha}{\op O}  \equiv 
\mat{ \trace{\excite{\alpha}{p}\op O } \;, & \trace{\decay{p}{\alpha}\op O} }
\;,
\ee
where the flip in the role of $\op e$ and $\op d$ in \eqref{TransitionBasisConjugateDef1} follows from the properties of excitation and decay operators
under hermitian-conjugate operation. The expansion of a suitable operator in the transition basis using the conjugate description reads: 
\be\lb{OperatorTransBasisConjugateExpDef1}
\op O = \sum_p\sint \dd \alpha \sbra{p\alpha}{\op O}\sket{p\alpha}{\cdot} = 
\dbraket{\op O}{\cdot} \;.
\ee

\subsubsection{Spectral decomposition of the response density in the transitions basis}

The Liouville equation \eqref{LiouvillianRhopomegaDef3} can be consistently written in the transition basis. Indeed, its r.h.s is explicitly written
as a commutator with $\dmnot$, while an inspection of the formal definition of the response density and of the Liouville superoperator evidences that
they both contain a commutator with $\dmnot$. Written in the basis of transitions Liouville equation becomes a linear matrix equation for the rank-2 
vectors of the coefficients \eqref{OperatorTransitionCoeffDef1} of the operators in its left and right side:
\be
\omega\mat{ \trace{\decay{\alpha}{p}\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\dm'(\omega)}} - 
\mat{ \trace{\decay{\alpha}{p}\Liouv\dm'(\omega)} \\ \trace{\excite{p}{\alpha}\Liouv\dm'(\omega)}} =
\mat{ \trace{\decay{\alpha}{p}\commutator{\op\Phi(\omega)}{\dmnot}} \\ \trace{\excite{p}{\alpha}\commutator{\op\Phi(\omega)}{\dmnot}}}
\;.\nn
\ee
The second addend in the l.h.s can be further analyzed by writing the transition basis expansion of $\dm'$ inside the traces. Performing this procedure
we can recast this quantity as the action of a 2-dimensional matrix on the rank-2 vector of the response density. So we have: 
\be\lb{LiouvilleTransBasisDef1}
\sum_q\sint\dd\beta \left[\omega\delta_q^p\delta(\alpha-\beta)- \optr{L}^{p \alpha}_{q\beta} \right]\sket{q\beta}{\dm'(\omega)} =
\sket{p\alpha}{\commutator{\op\Phi(\omega)}{\dmnot}} \;.
\ee
with:
\be\lb{LiouvilleMatrixTransBasisDef1}
\optr{L}^{p \alpha}_{q\beta} \equiv \mat{\trace{\decay{\alpha}{p}\, \Liouv\, \excite{q}{\beta}} &
\trace{\decay{\alpha}{p}\, \Liouv\, \decay{\beta}{q}} \\
\trace{\excite{p}{\alpha}\, \Liouv\, \excite{q}{\beta}} &
\trace{\excite{p}{\alpha}\, \Liouv\, \decay{\beta}{q}}} \;.
\ee
This formulation of the Liouville equation is the starting point for the derivation of the spectral decomposition of the response density operator
in terms of the eigenvectors if the Liouville operator. 
The first step in the achievement of this result is provided by the construction of the eigenvectors of the matrix \eqref{LiouvilleMatrixTransBasisDef1}.
Following the notation previously defined we introduce the rank-2 vectors $\sket{p\alpha}{E}$ and $\sbra{p\alpha}{\tilde E}$ that represent, by definition,
the right and left eigenvectors $\optr{L}$, respectively, so that:
\be\lb{LiouvMatrixEigenvectorDef1}
\sum_q \sint \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q \beta}{E} = E \sket{p\alpha}{E} \;, \quad
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \optr{L}^{p \alpha}_{q\beta} = E \sbra{q\beta}{\tilde E}\;.
\ee
In general, left and right eigenvectors do not coincide, but it is always possible to choose they normalization so that a orthonormality conditions among the
elements of left and right spaces holds, namely:
\be
\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E}\sket{p\alpha}{E'} \equiv \dbraket{\tilde E}{E'} = \delta(E-E') \;.
\ee
Once that the eigenvectors are known, the spectral decomposition of the resolvent of $\optr{L}$ can be provided:  
\be
\left[\left(\bbomega-\optr{L}\right)^{-1}\right]^{p\alpha}_{q\beta} =\sint\dd E
\frac{\sket{p\alpha}{E}\sbra{q\beta}{\tilde E}}{\bbomega-E} \;,
\ee
so equation \eqref{LiouvilleTransBasisDef1} can be inverted:
\be\lb{RhopExcitationTransBasisDef1}
\sket{p\alpha}{\dm'(\omega)} = \sint\dd E \sket{p\alpha}{E}
\frac{\sum_q \sint \dd \beta\sbra{q\beta}{\tilde E}
\sket{q\beta}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
that can be compactly rewritten by using the index-less notation:
\be\lb{RhopExcitationTransBasisDef2}
\sket{p\alpha}{\dm'(\omega)} = 
\sint\dd E \sket{p\alpha}{E}
\frac{\dbraket{\tilde E}{\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;.
\ee

@@

We introduce the explicit representation of the eigenvectors op $\optr{L}$:
\be
\sket{p\alpha}{E} \equiv \mat{X_E^{p\alpha} \\ Y_E^{p\alpha}} \;, \qq
\sbra{p\alpha}{\tilde E} \equiv \mat{\tilde X_{E\,p\alpha} & \tilde Y_{E\,p\alpha}} \;,
\ee
and on the basis of these elements we define the \emph{excitation operators}:
\begin{align}\lb{ExctitationOperatorsDef1}
\op E &\equiv \dbraket{\cdot}{E} = 
\sum_p \sint \dd \alpha \left(\excite{p}{\alpha} X_E^{p\alpha} + \decay{\alpha}{p} Y_E^{p\alpha}\right) 
\;,\nn \\
\op{\tilde E} &\equiv \dbraket{\tilde E}{\cdot} = \sum_p \sint \dd \alpha \left(\decay{\alpha}{p} \tilde X_{E\,p\alpha} + 
\excite{p}{\alpha}\tilde Y_{E\,p\alpha}  \right)\;.
\end{align}
Excitation operators are the right and left eigenvectors of the Liouville superoperator, as directly follows from their
definition. Indeed for $\op E$ we have:
\begin{align}
\Liouv \op E & = \sum_p \sint \dd \alpha \left(\trace{\decay{\alpha}{p} \Liouv \op E} \excite{p}{\alpha} + 
\decay{\alpha}{p} \trace{ \excite{p}{\alpha} \Liouv \op E}  \right) = \nn \\
&= \sum_p \sint  \dd \alpha \sbra{p\alpha}{\cdot} \sum_q \sint  \dd \beta \optr{L}^{p \alpha}_{q\beta} \sket{q\beta}{E} 
= E \sum_p \sint \dd \alpha \sbra{p\alpha}{\cdot} \sket{p\alpha}{E} = E \, \op E \;,	
\end{align}
whereas the left action on $\op{\tilde E}$ provides:
\begin{align}
 \op{\tilde E} \Liouv &=  \sum_p \sint \dd \alpha \left(
\trace{\op{\tilde E}\Liouv\excite{p}{\alpha}}\decay{\alpha}{p} +  
\trace{\op{\tilde E}\sop{\mathcal L}\decay{\alpha}{p}} \excite{p}{\alpha} \right) = 
\nn \\ &= \sum_p \sint  \dd \alpha  \sum_q \sint  \dd \beta
\sbra{q\beta}{\tilde E}\optr{L}^{q\beta}_{p\alpha}\sket{p\alpha}{\cdot} = 
E\sum_p \sint \dd \alpha \sbra{p\alpha}{\tilde E} \sket{p\alpha}{\cdot}  =
E \op{\tilde E} \;.
\end{align}
The introduction of the excitation operators allows to provided a complete operator version of equation \eqref{RhopExcitationTransBasisDef2}
\be\lb{RhopExcitationDef1}
\dm'(\omega) = 
\sint\dd E \op E
\frac{\trace{\op{\tilde E}\commutator{\op\Phi(\omega)}{\dmnot}}}{\bbomega-E} \;,
\ee
where the algebraic equivalence between the trace written in \eqref{RhopExcitationDef1} and the numerator of \eqref{RhopExcitationTransBasisDef2}
can be easily proved.
